My sincerest thanks to my academic advisor, Professor Shawn Blanton.
%
I am deeply grateful for his continued support and patience as I struggled through this long process.
%
He has given me valuable instruction not only in the technical aspects of this particular domain, but wider lessons on academic research and writing, while providing constant encouragement and a positive attitude.
%
This thesis would not have been possible without his guidance and advice.

I would like to thank the other members of my committee, Prof. Ken Mai, Prof. Diana Marculescu, and Prof. Subhasish Mitra, for finding time to share their knowledge and advice as part of my thesis committee.
%
Their suggestions, ideas, and feedback have helped me to focus the ideas we originally discussed during my thesis proposal into this final product.

Many thanks to the dedicated staff of the ECE department, including Adam Palko, Judy Bandola, Elaine Lawrence, Susan Farrington, Samantha Goldstein, and Nathan Snizaski.
%


I am sincerely grateful to all past and present ACTL students, including Jeff Nelson, Osei Poku, Yen-Tzu Lin, Wing-Chiu Tam, Xiaochun Yu, John Porche, Mitch Martin, Hongfei Wang, Cheng Xue, Xuanle Ren, Ben Niewenhuis, and Martyn Romanko - Your friendship and assistance during our years together was invaluable.
%
Special thanks to Ben Niewenhuis and Harsh Shrivastava for their research contributions.
%
Also thanks to other ECE grad students including Peter Klemperer, Adam Hartman, Kristen Dorsey, John Kelly, Craig Teegarden, Scott Fisk, Stephen Powell, Joey Fernandez, Mark McCartney, Peter Milder, and everyone else I am forgetting here - I sincerely enjoyed the many shared meals, cups of coffee, and camaraderie.
%
Many thanks to the ECE Graduate Organization for organizing social events and providing opportunities for volunteer service.
%
Further thanks to the ECE department softball team (The Gigahurtz), for getting the pale nerds outside in the fresh air once in a while.
%
Thanks to Dr. Yanjing Li for our many fruitful discussions of on-chip test, NBTI, ELF, and many other topics through the years.

To my parents Rick and Linda Beckler, who have always supported my wildest dreams and craziest plans, thank you for the never-ending support.
%
Thank you for all the encouragement through this whole process, your visits to Pittsburgh, and the excellent holidays away from school.

Finally, this dissertation is dedicated to my dearest wife, Anna.
%
Thank you for following me across the country and making a life together far from home.
%
Thank you for keeping me sane during this journey, and sticking it out with me until the end.
%
Thank you for your endless love and support, I could not have done this without you.

\sectionline

The work in this dissertation was supported by the Semiconductor Research Corporation through contract no. 1974.001, the National Science Foundation under contract no. 0903478, as well as the Lamme/Westinghouse and Bertucci Graduate Fellowships.

